---
title: "Dplyr vs Pandas"
author: "Miguel Angel Naranjo Mora + 97943"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reticulate)
```

Load the Iris dataset in R. and remove the levels of the class variable.

```{r}
library(dplyr)
iris_df <- iris
class(iris_df$Species)

iris_df$Species <- as.character(iris_df$Species)

head(iris_df)
```

Rename the columns to prevent confusions in Python (dot is python accessor)

```{r}
colnames(iris_df) <- sub("\\.","_",colnames(iris_df))
head(iris_df)
```

Load pandas. Use iris dataset from R in Python

```{python}
import pandas as pd
iris_python = pd.DataFrame(r.iris_df)
type(iris_python)
iris_python.head()
```

**Warning!**

In **Pandas** you can use the **filter()** method. This is confusing because the **filter()** function in **dplyr** is used to subset rows based on conditions and not columns! In **dplyr** we use the **select()** function instead.

```{r}
iris_df %>%
select(Sepal_Width, Petal_Width)%>%
head()
```

```{python}
#Pass columns as list using [[]]
iris_python[["Sepal_Width","Petal_Width"]]
#Or use Filter Function
iris_python.filter(items=['Sepal_Width','Petal_Width'])
```

Filter based on conditions

```{r}
iris_df %>% filter(Sepal_Width > 3.5 & Petal_Width < 0.3)
```

```{python}

#query API
iris_python.query("Sepal_Width > 3.5 & Petal_Width < 0.3")

#indexing
iris_python[(iris_python["Sepal_Width"] > 3.5) & (iris_python["Petal_Width"] < 0.3)]
```

Renaming a single column

```{r}
iris_df <- iris_df %>% rename(Class=Species)
head(iris_df)
```

```{python}
iris_python.rename(columns = {'Species': 'Class'}, inplace = True)
iris_python.head()
```

Renaming multiple columns based on condition

```{r}
iris_df <- iris_df %>%
rename_with(toupper,matches("length|width"))
head(iris_df)
```

```{python}
import re
pattern = re.compile(r".*(Length|Width)")
for col in iris_python.columns:
  if bool((pattern.match(col))):
    iris_python.rename(columns = {col: col.upper()}, inplace = True)
iris_python.head()
```

Altering cell values based on conditions

```{r}
iris_df <- iris_df %>%
mutate(Class = case_when(Class == 'setosa' ~ 0
                           ,Class == 'versicolor' ~ 1
                           ,Class == 'virginica' ~ 2))
head(iris_df)
```

```{python}
iris_python.loc[iris_python['Class'] == 'setosa', "Class"] = 0
iris_python.loc[iris_python['Class'] == 'versicolor', "Class"] = 1
iris_python.loc[iris_python['Class'] == 'virginica', "Class"] = 2
iris_python.head()
```
